---
title: "Lab 9"
author: "Caleb Jensen"
format: 
  html: 
    self-contained: true
editor: visual
---

## 1) Packages & Data

```{r}
#| output: false
library(broom)
library(tidyverse)
library(DT)
A_names <- read_csv(here::here("supporting_artifacts", "Lab 9", "StateNames_A.csv"))
```

## 2) Summary and Visualization

```{r}

Allison <- A_names |> 
  rename(Sex = Gender) |> 
  mutate(Count = replace_na(Count, 0)) |> 
  filter(Name == "Allison") |> 
  group_by(State, Year) |> 
  summarise(Count, Sex, .groups = "drop")

datatable(Allison, filter = 'top', class = 'cell-border stripe')

Allison <- Allison |> 
  filter(Sex == "F")

Allison |> 
  group_by(Year) |> 
  summarise(Count = sum(Count)) |> 
  ggplot(mapping = aes(x = Year, y = Count)) +
  geom_point() + 
  geom_path() + 
  labs(x = "", y = "", title = "Number of Babies Names Allison by Year") +
  theme(plot.title = element_text(hjust = 0.5))
```

## 3) Modeling The Number of Allisons

```{r}
ali_mod <- lm(Count ~ Year, data = Allison)
summary(ali_mod)
Allison |> 
  ggplot(mapping = aes(x = Year, y = Count)) +
    geom_point() + 
    stat_smooth(method = "lm") + 
    labs(x = "", y = "", title = "Plot of Allisons vs. Year With Regression Line") +
    theme(plot.title = element_text(hjust = 0.5))
```

The estimated regression equation is "Allisons = 3894.36 - 1.882 \* Year" This means the bigger year gets the fewer babies named Allison we expect.

## 4) Residuals in Allison Model

```{r}
ali_mod |> 
  ggplot(mapping = aes(x = .resid)) + 
  geom_density() +
  labs(x = "", y = "", title = "Density Plot of Residuals Predicting Count with Year") + 
  theme(plot.title = element_text(hjust = 0.5))
```

From this model we can say that, in general, the name "Allison" has gotten less popular through the years, however, it appears that correlation is not extremely strong and varies from state to state.

## 5) Babies Named "Allen" "Allan" or "Alan"

```{r}
Al_names <- data.frame(Name = c("Allan", "Alan", "Allen"), Sex = c("M", "M", "M"))

Allans <- A_names |> 
  rename(Sex = Gender) |> 
  semi_join(Al_names, by = c("Name", "Sex")) 

Allans |> 
  group_by(Year, Name) |> 
  summarise(Count = sum(Count)) |> 
  ggplot(mapping = aes(x = Year, y = Count, color = Name)) +
  geom_point() + 
  geom_path() +
  labs(x = "", y = "", title = "Count of Allen's by Spelling and Year") + 
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none") +
  annotate("text", x = 2010, y = 3000, label = "Alan") +
  annotate("text", x = 2010, y = 1100, label = "Allen") + 
  annotate("text", x = 2010, y = 550, label = "Allan") 
  

```

## 6) Spelling by State

```{r}
states <- data.frame(State = c("PA", "CA"))

Allans |> 
  filter(Year == 2000) |> 
  semi_join(states, by = "State") |> 
  pivot_wider(names_from = Name, 
              values_from = Count) |> 
  group_by(State) |> 
  mutate(Allan = replace_na(Allan, 0), 
         Total = sum(Allen, Allan, Alan),
         Alan = round(Alan/Total, 2),
         Allen = round(Allen/Total, 2), 
         Allan = round(Allan /Total, 2)) |> 
  datatable(class = "cell-border stripe")
  
```

Fewer Al named people where born in Pennsylvania overall and the proportions were somewhat different. While a large proportion of "Als" in both states were spelled "Alan" a even larger proportion were spelled "Allen" in Pennsylvania, making it much more common to have that spelling as a Pennsylvanian than as a Californian.
